---
title: "Assignment 06"
author: "Robert Wilbrand"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)
library(brms)
data(tulips, package = "rethinking")

# set global options
theme_set(theme_minimal())
knitr::opts_knit$set(root.dir = "C:/MSc GCG/2021_SoSe/ASM")
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
options(scipen = 8) # disables scientific notation up to 8 digits

# Wrapper function to set a specified seed prior to function execution
# useful because it leaves the seed outside the function unaffected
seedFunc <- function(seed, func, ...){
  set.seed(seed)
  func(...)
}
```

## 8H1.

Return to the data(tulips) example in the chapter. Now include the bed variable as a predictor in the interaction model. Don’t interact bed with the other predictors; just include it as a main effect. Note that bed is categorical. So to use it properly, you will need to either construct dummy variables or rather an index variable, as explained in Chapter 5.

```{r}
# rescale predictors, factorize bed
tulip <- tulips %>% 
  mutate(blooms_std = blooms/max(blooms),
         water_ctd  = water - mean(water),
         shade_ctd  = shade - mean(shade),
         across(bed, as.factor))
rm(tulips)
```

```{r results='hide'}
# Compile models with and without bed
tulipBedModel <-
  brm(data = tulip, 
      family = gaussian,
      blooms_std ~ 0 + bed + water_ctd + shade_ctd + water_ctd:shade_ctd,
      prior = c(#prior(normal(0.5, 0.25), class = Intercept),
                prior(normal(0, 0.25), class = b, coef = water_ctd),
                prior(normal(0, 0.25), class = b, coef = shade_ctd),
                prior(normal(0, 0.25), class = b, coef = "water_ctd:shade_ctd"),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 825)

tulipNoBedModel <-
  brm(data = tulip, 
      family = gaussian,
      blooms_std ~ 1 + water_ctd + shade_ctd + water_ctd:shade_ctd,
      prior = c(prior(normal(0.5, 0.25), class = Intercept),
                prior(normal(0, 0.25), class = b, coef = water_ctd),
                prior(normal(0, 0.25), class = b, coef = shade_ctd),
                prior(normal(0, 0.25), class = b, coef = "water_ctd:shade_ctd"),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 825)
```

```{r}
# Add criteria
tulipBedModel <- add_criterion(tulipBedModel, criterion = "waic")
tulipNoBedModel <- add_criterion(tulipNoBedModel, criterion = "waic")

# Show plot and summary
plot(tulipBedModel)
summary(tulipBedModel)
```

## 8H2.

Use WAIC to compare the model from 8H1 to a model that omits bed. What do you infer from this comparison? Can you reconcile the WAIC results with the posterior distribution of the bed coefficients?

```{r}
# Compare models
comparison <- loo_compare(tulipBedModel, tulipNoBedModel, criterion = "waic")
print(comparison, simplify = FALSE)

# Extract and plot posterior samples
psamp <- posterior_samples(tulipBedModel)[,1:3] %>% 
  pivot_longer(cols = everything())
ggplot(psamp) +
  geom_density(mapping = aes(x = value, group = name, col = name, fill = name),
               alpha = 0.3) +
  theme_bw()
```

Adding bed to the model improves its predictive performance, but with great uncertainty. The distribution plots show that bed A is separable from B and C, but B and C have very similar distributions. That serves to explain why the difference in ELPD is less than half of the standard error of this difference.
